HW 2: Energy Demand and CO2 Emissions

### 2022-09-28 HW2/HW2.Rmd
update hw2

library(tidyverse)
library(gridExtra)
library(kableExtra)
```

@@

		powerraw = dpowerraw + apowerraw + rpowerraw + gpowerraw,
		power = ifelse(powerraw < 0, 0, powerraw)
		) %>% 
	select(-c(`Time (seconds)`, `Speed (mph)`, `Grade (rise/run)`)) %>% 
	relocate(time, v)
```

@@

A graph of engine power along with tractive power is given in Figure \ref{powercomp}. I am using only from `r timemin %>% format(scientific=FALSE)` to `r timemax %>% format(scientific=FALSE)` seconds for clarity of the graph.

```{r, fig.height=4, fig.cap="Comparison of engine and tractive power.\\label{powercomp}"}
powercomp <- tp %>% 
	filter(time <= timemax, time >= timemin) %>% 
	transmute(time, power, engine = `Engine Power (kW)`,
						diff = power - engine) %>% 
	pivot_longer(-time) %>% 
	mutate(name = as.factor(name))

@@

powercomp %>% 
	ggplot(aes(x = time, y = value, color = name)) +
	geom_line() +
	labs(x = "Time (seconds)",
			 y = "Power (kW)") +
	scale_color_discrete(name = "", labels = c(
		diff = "Engine - Tractive", engine = "Engine Power", power = "Tractive Power")) +
	theme_minimal() + 
  theme(legend.position = 'bottom')
```

The data generally seem to make sense, as the engine power is often more than the tractive power, which is expected due to losses. Though I am not sure why engine power sometimes exceeds tractive power; it appears to have something to do with the spikes in tractive power, which are perhaps due to some time gaps.

## 2.3
The total power required to move the vehicle for this drive cycle is `r sum(tp$power, na.rm=TRUE) %>% round() %>% format(scientific=FALSE)` kW.

@@

Assuming the "negative" tractive power would be used for regenerative braking, an additional `r (regen*hours) %>% round() %>% format(scientific=FALSE)` kWh would be available (`r regen %>% round() %>% format(scientific=FALSE)` Kw over `r hours %>% round(2)` hours). If 50% of that power was recaptured, then the additional tractive power needed would be reduced to `r newpower %>% round() %>% format(scientific=FALSE)` kW, a reduction of `r ((oldpower - newpower)/oldpower * 100) %>% round(2)`%.

>>>>

library(tidyverse)
library(gridExtra)
library(kableExtra)

options(scipen = 999)
```

@@

		powerraw = dpowerraw + apowerraw + rpowerraw + gpowerraw,
		power = ifelse(powerraw < 0, 0, powerraw),

		dist = v*(time - lag(time))
		) %>% 
	select(-c(`Time (seconds)`, `Speed (mph)`, `Grade (rise/run)`)) %>% 
	relocate(time, v)
```

@@

A graph of engine power along with tractive power is given in Figure \ref{powercomp}. I am using only from `r timemin` to `r timemax` seconds for clarity of the graph.

```{r, fig.height=4, fig.cap="Comparison of engine and tractive power.\\label{powercomp}"}
powercomp <- tp %>% 
	filter(time <= timemax, time >= timemin) %>% 
	transmute(time, power, engine = `Engine Power (kW)`,
						diff = engine - power) %>% 
	pivot_longer(-time) %>% 
	mutate(name = as.factor(name))

@@

powercomp %>% 
	ggplot(aes(x = time, y = value, color = name)) +
	geom_line() +
	labs(x = "Time (seconds)",
			 y = "Power (kW)") +
	scale_color_discrete(name = "", labels = c(
		diff = "Tractive minus Engine", engine = "Engine Power", power = "Tractive Power")) +
	theme_minimal() + 
  theme(legend.position = 'bottom')
```

The data generally seem to make sense, as the engine power is often more than the tractive power, which is expected due to losses. Though I am not sure why tractive power sometimes exceeds engine power; it appears to have something to do with the spikes in tractive power, which are perhaps due to some time gaps.

## 2.3
The total power required to move the vehicle for this drive cycle is `r sum(tp$power, na.rm=TRUE) %>% round()` kW.

@@

Assuming the "negative" tractive power would be used for regenerative braking, an additional `r (regen*hours) %>% round()` kWh would be available (`r regen %>% round()` kW over `r hours %>% round(2)` hours). If 50% of that power was recaptured, then the additional tractive power needed would be reduced to `r newpower %>% round()` kW, a reduction of `r ((oldpower - newpower)/oldpower * 100) %>% round(2)`%.

## 2.6
```{r}
enginemove <- tp %>% 
	mutate(newengine = `Engine Power (kW)` - 7,
				 newengine = ifelse(newengine < 0, 0, newengine))

moveenergy <- sum(enginemove$newengine, na.rm = TRUE)*hours
```

Assuming an auxiliary power load of 7 kW, the engine power used to move the vehicle is $P_{Engine} - 7$ kW, but not less than 0 kW. Finding the total energy expended by the engine attempting to move the vehicle gives `r moveenergy %>% round()` kWh. The total energy actually required to move the vehicle is `r (oldpower*hours) %>% round()` kWh, so the efficiency of the drivetrain is `r (oldpower*hours/moveenergy * 100) %>% round(1)`%.

## 2.7
Low-sulfur diesel has an energy content of 128,488 Btu/gal according to the AFDC. This is equivalent to `r (energy <- 37.656)` kWh/gal. Assuming an engine efficiency of `r (eff <- 0.3)`, the total fuel needed for this drive cycle is $F[\text{gal}] = \frac{`r (oldpower*hours) %>% round()`[\text{kWh}]}{`r energy`[\text{kWh/gal}]}/`r eff` = `r (fuel <- oldpower*hours/energy/eff) %>% round()`\ \text{gal}$. The total distance in this drive cycle is `r (totdist <- sum(tp$dist, na.rm = TRUE)) %>% round()` meters or `r (totdist * .0006213712) %>% round(1)` miles, giving
